### 2023-09-14 proofs/main.tex
Add rule labels
WIP: add comments

\renewcommand{\ottdruleONVarInName}[0]{(Var$_{\in}^-$)}
\renewcommand{\ottdruleONVarNInName}[0]{(Var$_{\notin}^-$)}
\renewcommand{\ottdruleONUVarName}[0]{(UVar$^-$)}
\renewcommand{\ottdruleOShiftUName}[0]{($\uparrow$)}
\renewcommand{\ottdruleOArrowName}[0]{($\rightarrow$)}
\renewcommand{\ottdruleOForallName}[0]{($\forall$)}

@@

\renewcommand{\ottdruleOPVarInName}[0]{(Var$_{\in}^+$)}
\renewcommand{\ottdruleOPVarNInName}[0]{(Var$_{\notin}^+$)}
\renewcommand{\ottdruleOPUVarName}[0]{(UVar$^+$)}
\renewcommand{\ottdruleOShiftDName}[0]{($\downarrow$)}
\renewcommand{\ottdruleOExistsName}[0]{($\exists$)}

% nf(N) = M
\renewcommand{\ottdruleNrmNVarName}[0]{(Var$^-$)}
\renewcommand{\ottdruleNrmNUVarName}[0]{(UVar$^-$)}
\renewcommand{\ottdruleNrmShiftUName}[0]{($\uparrow$)}
\renewcommand{\ottdruleNrmArrowName}[0]{($\rightarrow$)}
\renewcommand{\ottdruleNrmForallName}[0]{($\forall$)}

% nf(P) = Q
\renewcommand{\ottdruleNrmPVarName}[0]{(Var$^+$)}
\renewcommand{\ottdruleNrmPUVarName}[0]{(UVar$^+$)}
\renewcommand{\ottdruleNrmShiftDName}[0]{($\downarrow$)}
\renewcommand{\ottdruleNrmExistsName}[0]{($\exists$)}

@@

\renewcommand{\ottdruleEOneNVarName}[0]{(Var$^-$$^{\eqEOne}$)}
\renewcommand{\ottdruleEOneShiftUName}[0]{($\uparrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneArrowName}[0]{($\rightarrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneForallName}[0]{($\forall^{\eqEOne}$)}

% P ≈ Q
\renewcommand{\ottdruleEOnePVarName}[0]{(Var$^+$$^{\eqEOne}$)}
\renewcommand{\ottdruleEOneShiftDName}[0]{($\downarrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneExistsName}[0]{($\exists^{\eqEOne}$)}

@@

\renewcommand{\ottdruleDOneNVarName}[0]{(Var$^-$$^{\subDOne}$)}
\renewcommand{\ottdruleDOneShiftUName}[0]{($\uparrow^{\subDOne}$)}
\renewcommand{\ottdruleDOneArrowName}[0]{($\rightarrow^{\subDOne}$)}
\renewcommand{\ottdruleDOneForallName}[0]{($\forall^{\subDOne}$)}

% G ⊢ P ≥1 Q
\renewcommand{\ottdruleDOnePVarName}[0]{(Var$^+$$^{\supDOne}$)}
\renewcommand{\ottdruleDOneShiftDName}[0]{($\downarrow^{\supDOne}$)}
\renewcommand{\ottdruleDOneExistsName}[0]{($\exists^{\supDOne}$)}

% G ⊢ N ≈1 M
\renewcommand{\ottdruleDOneNDefName}[0]{($\eqDOne^{-}$)}

% G ⊢ P ≈1 Q
\renewcommand{\ottdruleDOnePDefName}[0]{($\eqDOne^{+}$)}

@@

% G ; Θ ⊨ uN ≤ iM ⫤ us
\renewcommand{\ottdruleANVarName}[0]{(Var$^-$$^{\subA}$)}
\renewcommand{\ottdruleAShiftUName}[0]{($\uparrow^{\subA}$)}
\renewcommand{\ottdruleAArrowName}[0]{($\rightarrow^{\subA}$)}
\renewcommand{\ottdruleAForallName}[0]{($\forall^{\subA}$)}

% G ; Θ ⊨ iP ≥ uQ ⫤ us
\renewcommand{\ottdruleAPVarName}[0]{(Var$^+$$^{\supA}$)}
\renewcommand{\ottdruleAShiftDName}[0]{($\downarrow^{\supA}$)}
\renewcommand{\ottdruleAExistsName}[0]{($\exists^{\supA}$)}
\renewcommand{\ottdruleAPUVarName}[0]{(UVar$^{\supA}$)}

@@

% Γ ; Θ ⊨ uN ≈u iM ⫤ UC 
\renewcommand{\ottdruleUNVarName}[0]{(Var$^{-[[≈u]]}$)}
\renewcommand{\ottdruleUShiftUName}[0]{($\uparrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUArrowName}[0]{($\rightarrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUForallName}[0]{($\forall^{[[≈u]]}$)}
\renewcommand{\ottdruleUNUVarName}[0]{(UVar$^{-[[≈u]]}$)}

% Γ ; Θ ⊨ uP ≈u iQ ⫤ UC 
\renewcommand{\ottdruleUPVarName}[0]{(Var$^{+[[≈u]]}$)}
\renewcommand{\ottdruleUShiftDName}[0]{($\downarrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUExistsName}[0]{($\exists^{[[≈u]]}$)}
\renewcommand{\ottdruleUPUVarName}[0]{(UVar$^{+[[≈u]]}$)}

% G ⊨ iP1 ≈au iP2 ⫤ ( Ξ , uQ , aus1 , aus2 )
\renewcommand{\ottdruleAUPVarName}[0]{(Var$^{+[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftDName}[0]{($\downarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUExistsName}[0]{($\exists^{[[≈au]]}$)}

% G ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )
\renewcommand{\ottdruleAUNVarName}[0]{(Var$^{-[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftUName}[0]{($\uparrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUForallName}[0]{($\forall^{[[≈au]]}$)}
\renewcommand{\ottdruleAUArrowName}[0]{($\rightarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUAUName}[0]{(AU$^{-}$)}

@@

\ottgrammartabular{
  \ottiP\ottinterrule
  \ottiN\ottinterrule
}

\subsection{Equalities}
For simplicity, assume alpha-equivalent terms equal. 
This way, we assume that substitutions do not capture bound variables.
Besides, we equate
$[[∀pas.∀pbs.iN]]$ with $[[∀pas,pbs.iN]]$, 
as well as $[[∃nas.∃nbs.iP]]$ with $[[∃nas,nbs.iP]]$,
and lift these equations transitively and congruently 
to the whole system.

@@

\hfill\\
\ottdefnsWFT

@@

\begin{definition}[Substitution Application]
  Substitution application (denoted as $[[ [σ]iP ]]$ or $[[ [σ]iN ]]$) 
  is defined congruently as follows:
  \begin{itemize}
    \item $[[ [σ]α⁺ ]] = [[σ]] ([[α⁺]])$;
    \item $[[ [σ]α⁻ ]] = [[σ]] ([[α⁻]])$;
    \item $[[ [σ]↓iN ]] = [[↓[σ]iN]]$;
    \item $[[ [σ]↑iP ]] = [[↑[σ]iP]]$;
    \item $[[ [σ]∃nas.iQ ]] = [[∃nas.[σ]iQ]]$, 
    \item $[[ [σ]∀pas.iN ]] = [[∀pas.[σ]iN]]$(here we assume that $[[nas]]$ and $[[pas]]$ are lists of fresh variables, 
      that is the variable capture never happens);
    \item $[[ [σ](iP → iN) ]] = [[ [σ]iP → [σ]iN ]]$.
  \end{itemize}
\end{definition}

@@

Sometimes it is convenient to construct substitution 
explicitly mapping each variable from a list (or a set)
to a type. Such substitutions are denoted as $[[iPs / pas]]$
and $[[iNs / nas]]$, where $[[iPs]]$ and $[[iNs]]$ are lists of 
the corresponding types.
\begin{definition}[Explicit Substitution]
  \hfill
  \begin{itemize}
    \item [$-$]
      Suppose that $[[nas]]$ is a list of negative type variables,
      and $[[iNs]]$ is a list of negative types of the same length 
      Then $[[iNs / nas]]$ denotes a substitution such that 
      \begin{enumerate}
        \item for $[[αi⁺ ∊ {nas}]]$, $[[ [iNs / nas] αi⁺]] = [[iNi]]$;
        \item for $[[β⁺ ∉ {nas}]]$, $[[ [iNs / nas] β⁺]] = [[β⁺]]$.
      \end{enumerate}
    \item [$+$]
      Positive explicit substitution $[[iPs / pas]]$
      is defined symmetrically.
  \end{itemize}
\end{definition}

\subsection{Declarative Subtyping}
\ottdefnsDOne

\subsection{Declarative Equivalence}
\ottdefnsEOne

@@

\begin{definition}[Algorithmic Types]
  \hfill \\
\ottgrammartabular{
  \ottuP\ottinterrule
  \ottuN\ottinterrule
}
\end{definition}

\subsection{Fresh Variable Selection}
\label{sec:fresh-selection}
Both the subtyping and the type inference algorithm
rely on the ability to select fresh, unused variables.
In the inference rules, it is indicated as 
$[[{varset} are fresh]]$. 
We assume that the selection subroutine always succeeds and is 
deterministic. In other words, whenever it is called in 
an algorithmic inference rule, it returns the same result, 
uniquely determined by the input of this rule.

\subsection{Variable Algorithmization}

In several places of our algorithm, in particular, during anti-unification,
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.

\begin{definition}[Variable Algorithmization]
  Suppose that $[[nas]]$ is a list of negative type variables
  and $[[nuas]]$ is a list of negative algorithmic variables of the same length. 
  Then $[[ nuas/nas ]]$ is a substitution-like procedure turning each $[[αi⁻ ∊ {nas}]]$ 
  into $[[αî⁻ ∊ {nuas}]]$.
\end{definition}

Conversely, we can turn algorithmic type variables
into declarative type variables via \emph{dealgorithmization}.

\begin{definition}[Variable Dealgorithmization]
  Suppose that $[[nuas]]$ is a list of negative algorithmic variables
  and $[[nas]]$ is a list of negative type variables of the same length. 
  Then $[[ nas/nuas ]]$ is a substitution-like procedure turning each
  $[[αî⁻ ∊ {nuas}]]$ into $[[αi⁻ ∊ {nas}]]$.
\end{definition}

@@

\begin{definition}[Algorithmic Type Context]
  \hfill \\
  Algorithmic type context $[[Ξ]]$ is represented by a set of 
  \emph{algorithmic} type variables ($[[α̂⁺]]$, $[[α̂⁻]]$, $[[β̂⁺]]$, \dots).
  The concatenation $[[Ξ1, Ξ2]]$ means the union of two contexts $[[Ξ1 ∪ Ξ2]]$.
\end{definition}

@@

\hfill\\
\ottdefnsWFAT

Algorithmic Type Context are used in the unification algorithm.
In the subtyping algorithm, 
the context needs to remember additional information.
In the subtyping context, each algorithmic variable is associated with a
context it must be instantiated in 
(i.e. the context in which the type replacing the variable must be well-formed).
This association is represented by \emph{algorithmic subtyping context} $[[Θ]]$.
\begin{definition}[Algorithmic Subtyping Context]
  \hfill \\
  Algorithmic Subtyping Context $[[Θ]]$ is represented by a set of 
  entries of form $[[ α̂⁺[Γ] ]]$ and $[[ α̂⁻[Γ] ]]$,
  where $[[α̂⁺]]$ and $[[α̂⁻]]$ are algorithmic variables,
  and $[[Γ]]$ is a context in which they must be instantiated.
  We assume that no two entries associating the same variable
  appear in $[[Θ]]$.

  $[[dom(Θ)]]$ denotes the set of variables appearing in $[[Θ]]$:
  $[[dom(Θ)]] = \{ [[α̂±]] \mid [[α̂±[Γ] ]] \in [[Θ]] \}$.

  If $[[ α̂±[Γ] ]] \in [[Θ]]$, we denote $[[Γ]]$ as $[[Θ(α̂±)]]$.
\end{definition}

@@

Substitution that operates on algorithmic type variables is denoted as
$[[uσ]]$. It is defined as a total function from algorithmic 
type variables to non-algorithmic types, preserving the polarity.

@@

\subsection{Normalization}

\subsubsection{Ordering}
\ottdefnsOrder

\subsubsection{Quantifier Normalization}
\ottdefnsNrm

We also define normalization of a substitution pointwise:
\begin{definition}[Substitution Normalization]
  For a substitution $[[σ]]$, we define $[[nf(σ)]]$
  as a substitution that maps $[[α±]]$ into $[[nf([σ]α±)]]$.
\end{definition}

>>>>

\renewcommand{\ottdruleONVarInName}[0]{(Var$_{-\in}^{\text{Ord}}$)}
\renewcommand{\ottdruleONVarNInName}[0]{(Var$_{-\notin}^{\text{Ord}}$)}
\renewcommand{\ottdruleONUVarName}[0]{(UVar$_{-} ^{\text{Ord}}$)}
\renewcommand{\ottdruleOShiftUName}[0]{($\uparrow^{\text{Ord}}$)}
\renewcommand{\ottdruleOArrowName}[0]{($\rightarrow^{\text{Ord}}$)}
\renewcommand{\ottdruleOForallName}[0]{($\forall^{\text{Ord}}$)}

@@

\renewcommand{\ottdruleOPVarInName}[0]{(Var$_{+\in}^{\text{Ord}}$)}
\renewcommand{\ottdruleOPVarNInName}[0]{(Var$_{+\notin}^{\text{Ord}}$)}
\renewcommand{\ottdruleOPUVarName}[0]{(UVar$_{+}^{\text{Ord}}$)}
\renewcommand{\ottdruleOShiftDName}[0]{($\downarrow^{\text{Ord}}$)}
\renewcommand{\ottdruleOExistsName}[0]{($\exists^{\text{Ord}}$)}

% nf(N) = M
\renewcommand{\ottdruleNrmNVarName}[0]{(Var$_{-}^{\text{nf}}$)}
\renewcommand{\ottdruleNrmNUVarName}[0]{(UVar$_{-}^{\text{nf}}$)}
\renewcommand{\ottdruleNrmShiftUName}[0]{($\uparrow^{\text{nf}}$)}
\renewcommand{\ottdruleNrmArrowName}[0]{($\rightarrow^{\text{nf}}$)}
\renewcommand{\ottdruleNrmForallName}[0]{($\forall^{\text{nf}}$)}

% nf(P) = Q
\renewcommand{\ottdruleNrmPVarName}[0]{(Var$_{-}^{\text{nf}}$)}
\renewcommand{\ottdruleNrmPUVarName}[0]{(UVar$_{-}^{\text{nf}}$)}
\renewcommand{\ottdruleNrmShiftDName}[0]{($\downarrow^{\text{nf}}$)}
\renewcommand{\ottdruleNrmExistsName}[0]{($\exists^{\text{nf}}$)}

@@

\renewcommand{\ottdruleEOneNVarName}[0]{(Var$_-^{\eqEOne}$)}
\renewcommand{\ottdruleEOneShiftUName}[0]{($\uparrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneArrowName}[0]{($\rightarrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneForallName}[0]{($\forall^{\eqEOne}$)}

% P ≈ Q
\renewcommand{\ottdruleEOnePVarName}[0]{(Var$_+^{\eqEOne}$)}
\renewcommand{\ottdruleEOneShiftDName}[0]{($\downarrow^{\eqEOne}$)}
\renewcommand{\ottdruleEOneExistsName}[0]{($\exists^{\eqEOne}$)}

@@

\renewcommand{\ottdruleDOneNVarName}[0]{(Var$_-^{\subDOne}$)}
\renewcommand{\ottdruleDOneShiftUName}[0]{($\uparrow^{\subDOne}$)}
\renewcommand{\ottdruleDOneArrowName}[0]{($\rightarrow^{\subDOne}$)}
\renewcommand{\ottdruleDOneForallName}[0]{($\forall^{\subDOne}$)}

% G ⊢ P ≥1 Q
\renewcommand{\ottdruleDOnePVarName}[0]{(Var$_+^{\supDOne}$)}
\renewcommand{\ottdruleDOneShiftDName}[0]{($\downarrow^{\supDOne}$)}
\renewcommand{\ottdruleDOneExistsName}[0]{($\exists^{\supDOne}$)}

% G ⊢ N ≈1 M
\renewcommand{\ottdruleDOneNDefName}[0]{($\eqDOne_{-}$)}

% G ⊢ P ≈1 Q
\renewcommand{\ottdruleDOnePDefName}[0]{($\eqDOne_{+}$)}

@@

% G ; Θ ⊨ uN ≤ iM ⫤ us
\renewcommand{\ottdruleANVarName}[0]{(Var$_-^{\subA}$)}
\renewcommand{\ottdruleAShiftUName}[0]{($\uparrow^{\subA}$)}
\renewcommand{\ottdruleAArrowName}[0]{($\rightarrow^{\subA}$)}
\renewcommand{\ottdruleAForallName}[0]{($\forall^{\subA}$)}

% G ; Θ ⊨ iP ≥ uQ ⫤ us
\renewcommand{\ottdruleAPVarName}[0]{(Var$_+^{\supA}$)}
\renewcommand{\ottdruleAShiftDName}[0]{($\downarrow^{\supA}$)}
\renewcommand{\ottdruleAExistsName}[0]{($\exists^{\supA}$)}
\renewcommand{\ottdruleAPUVarName}[0]{(UVar$^{\supA}$)}

@@

% Γ ; Θ ⊨ uN ≈u iM ⫤ UC 
\renewcommand{\ottdruleUNVarName}[0]{(Var$_{-}^{[[≈u]]}$)}
\renewcommand{\ottdruleUShiftUName}[0]{($\uparrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUArrowName}[0]{($\rightarrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUForallName}[0]{($\forall^{[[≈u]]}$)}
\renewcommand{\ottdruleUNUVarName}[0]{(UVar$_{-}^{[[≈u]]}$)}

% Γ ; Θ ⊨ uP ≈u iQ ⫤ UC 
\renewcommand{\ottdruleUPVarName}[0]{(Var$_{+}^{[[≈u]]}$)}
\renewcommand{\ottdruleUShiftDName}[0]{($\downarrow^{[[≈u]]}$)}
\renewcommand{\ottdruleUExistsName}[0]{($\exists^{[[≈u]]}$)}
\renewcommand{\ottdruleUPUVarName}[0]{(UVar$_{+}^{[[≈u]]}$)}

% G ⊨ iP1 ≈au iP2 ⫤ ( Ξ , uQ , aus1 , aus2 )
\renewcommand{\ottdruleAUPVarName}[0]{(Var$_{+}^{[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftDName}[0]{($\downarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUExistsName}[0]{($\exists^{[[≈au]]}$)}

% G ⊨ iN1 ≈au iN2 ⫤ ( Ξ , uM , aus1 , aus2 )
\renewcommand{\ottdruleAUNVarName}[0]{(Var$_{-}^{[[≈au]]}$)}
\renewcommand{\ottdruleAUShiftUName}[0]{($\uparrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUForallName}[0]{($\forall^{[[≈au]]}$)}
\renewcommand{\ottdruleAUArrowName}[0]{($\rightarrow^{[[≈au]]}$)}
\renewcommand{\ottdruleAUAUName}[0]{(AU$_{-}$)}

% Γ ⊢ iN 
\renewcommand{\ottdruleWFTNVarName}[0]{(Var$_{-}^{\text{WF}}$)}
\renewcommand{\ottdruleWFTShiftUName}[0]{($\uparrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFTArrowName}[0]{($\rightarrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFTForallName}[0]{($\forall^{\text{WF}}$)}

% Γ ⊢ iP
\renewcommand{\ottdruleWFTPVarName}[0]{(Var$_{+}^{\text{WF}}$)}
\renewcommand{\ottdruleWFTShiftDName}[0]{($\downarrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFTExistsName}[0]{($\exists^{\text{WF}}$)}

% Γ ; Ξ ⊢ uN
\renewcommand{\ottdruleWFATNVarName}[0]{(Var$_{-}^{\text{WF}}$)}
\renewcommand{\ottdruleWFATNUVarName}[0]{(UVar$_{-}^{\text{WF}}$)}
\renewcommand{\ottdruleWFATShiftUName}[0]{($\uparrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFATArrowName}[0]{($\rightarrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFATForallName}[0]{($\forall^{\text{WF}}$)}

% Γ ; Ξ ⊢ uP
\renewcommand{\ottdruleWFATPVarName}[0]{(Var$_{+}^{\text{WF}}$)}
\renewcommand{\ottdruleWFATPUVarName}[0]{(UVar$_{+}^{\text{WF}}$)}
\renewcommand{\ottdruleWFATShiftDName}[0]{($\downarrow^{\text{WF}}$)}
\renewcommand{\ottdruleWFATExistsName}[0]{($\exists^{\text{WF}}$)}

% Γ ⊢ iP : scE
\renewcommand{\ottdruleSATSCESupName}[0]{($[[:≥]]_{+}^\text{sat}$)}
\renewcommand{\ottdruleSATSCEPEqName}[0]{($[[:≈]]_{+}^\text{sat}$)}

% Γ ⊢ iN : scE
\renewcommand{\ottdruleSATSCENEqName}[0]{($[[:≈]]_{-}^\text{sat}$)}

% Γ ; Φ ⊢ v : iP 
\renewcommand{\ottdruleDTVarName}[0]{(Var$^{\text{inf}}$)}
\renewcommand{\ottdruleDTThunkName}[0]{($\{\}^{\text{inf}}$)}
\renewcommand{\ottdruleDTPAnnotName}[0]{(ann$_+^{\text{inf}}$)}
\renewcommand{\ottdruleDTPEquivName}[0]{($[[≈]]_+^{\text{inf}}$)}

% Γ ; Φ ⊢ c : iN 
\renewcommand{\ottdruleDTtLamName}[0]{($\lambda^{\text{inf}}$)}
\renewcommand{\ottdruleDTTLamName}[0]{($\Lambda^{\text{inf}}$)}
\renewcommand{\ottdruleDTReturnName}[0]{(ret$^{\text{inf}}$)}
\renewcommand{\ottdruleDTVarLetName}[0]{(let$^{\text{inf}}$)}
\renewcommand{\ottdruleDTAppLetName}[0]{(let$_@^{\text{inf}}$)}
\renewcommand{\ottdruleDTAppLetAnnName}[0]{(let$_{:@}^{\text{inf}}$)}
\renewcommand{\ottdruleDTUnpackName}[0]{(let$_{\exists}^{\text{inf}}$)}
\renewcommand{\ottdruleDTNAnnotName}[0]{(ann$_-^{\text{inf}}$)}
\renewcommand{\ottdruleDTNEquivName}[0]{($[[≈]]_-^{\text{inf}}$)}

% Γ ; Φ ⊢ iN ● args ⇒> iM 
\renewcommand{\ottdruleDTEmptyAppName}[0]{($\emptyset_{[[●]][[⇒>]]}^{\text{inf}}$)}
\renewcommand{\ottdruleDTArrowAppName}[0]{($\rightarrow_{[[●]][[⇒>]]}^{\text{inf}}$)}
\renewcommand{\ottdruleDTForallAppName}[0]{($\forall_{[[●]][[⇒>]]}^{\text{inf}}$)}

% Γ ; Φ ⊨ v : iP  
\renewcommand{\ottdruleATVarName}[0]{(Var$^{\text{inf}}$)}
\renewcommand{\ottdruleATThunkName}[0]{($\{\}^{\text{inf}}$)}
\renewcommand{\ottdruleATPAnnotName}[0]{(ann$_+^{\text{inf}}$)}

% Γ ; Φ ⊨ c : iN 
\renewcommand{\ottdruleATtLamName}[0]{($\lambda^{\text{inf}}$)}
\renewcommand{\ottdruleATTLamName}[0]{($\Lambda^{\text{inf}}$)}
\renewcommand{\ottdruleATReturnName}[0]{(ret$^{\text{inf}}$)}
\renewcommand{\ottdruleATVarLetName}[0]{(let$^{\text{inf}}$)}
\renewcommand{\ottdruleATAppLetName}[0]{(let$_@^{\text{inf}}$)}
\renewcommand{\ottdruleATAppLetAnnName}[0]{(let$_{:@}^{\text{inf}}$)}
\renewcommand{\ottdruleATUnpackName}[0]{(let$_{\exists}^{\text{inf}}$)}
\renewcommand{\ottdruleATNAnnotName}[0]{(ann$_-^{\text{inf}}$)}

% Γ ; Φ ; Θ1 ⊨ uN ● args ⇒> uM ⫤ Θ2 ; SC 
\renewcommand{\ottdruleATEmptyAppName}[0]{($\emptyset_{[[●]][[⇒>]]}^{\text{inf}}$)}
\renewcommand{\ottdruleATArrowAppName}[0]{($\rightarrow_{[[●]][[⇒>]]}^{\text{inf}}$)}
\renewcommand{\ottdruleATForallAppName}[0]{($\forall_{[[●]][[⇒>]]}^{\text{inf}}$)}

% scE1 singular with iP
\renewcommand{\ottdruleSINGPEqName}[0]{($[[≈]]_{+}^{\text{sing}}$)}
\renewcommand{\ottdruleSINGSupVarName}[0]{($[[:≥]]\alpha^{\text{sing}}$)}
\renewcommand{\ottdruleSINGSupShiftName}[0]{($[[:≥]][[↓]]^{\text{sing}}$)}

% scE1 singular with iN
\renewcommand{\ottdruleSINGNEqName}[0]{($[[≈]]_{-}^{\text{sing}}$)}

@@

% \ottgrammartabular{
%   \ottiP\ottinterrule
%   \ottiN\ottinterrule
% }

\begin{definition}[Declarative Types]
  \hfill
  \begin{multicols}{2}
    \ottgrammartabular{
      \ottiN\ottinterrule
    }

    \ottgrammartabular{
      \ottiP\ottinterrule
    }
    \columnbreak
  \end{multicols}
\end{definition}

\subsection{Equalities}
For simplicity, we assume alpha-equivalent terms equal.
This way, we assume that substitutions do not capture bound variables.
Besides, we equate
$[[∀pas.∀pbs.iN]]$ with $[[∀pas,pbs.iN]]$, 
as well as $[[∃nas.∃nbs.iP]]$ with $[[∃nas,nbs.iP]]$,
and lift these equations transitively and congruently 
to the whole system.

@@

Notice that checking the well-formedness of a type
is an \emph{algorithmic} procedure, in which 
both the context and the type are considered inputs.
In other words, it is syntax-directed and mode-correct 
(according to \cite{dunfieldBidirectionalTyping2020}), 
which means that

\begin{algorithm}[Type Well-formedness]
  \hfill

  \begin{multicols}{2}
  \ottdefnWFTNiWF{}
  \columnbreak

  \ottdefnWFTPiWF{}
  \end{multicols}

\end{algorithm}

@@

\begin{algorithm}[Substitution Application]
  Substitution application is denoted as $[[ [σ]iP ]]$ and $[[ [σ]iN ]]$.
  It is defined naturally as follows:
    \begin{multicols}{2}
      \begin{itemize}
        \item[] $[[ [σ]α⁺ ]] = [[σ]] ([[α⁺]])$;
        \item[] $[[ [σ]α⁻ ]] = [[σ]] ([[α⁻]])$;
        \item[] $[[ [σ]↓iN ]] = [[↓[σ]iN]]$;
        \item[] $[[ [σ]↑iP ]] = [[↑[σ]iP]]$;
        \item[] $[[ [σ](iP → iN) ]] = [[ [σ]iP → [σ]iN ]]$.
        \item[] $[[ [σ]∃nas.iQ ]] = [[∃nas.[σ]iQ]]$, 
        \item[] $[[ [σ]∀pas.iN ]] = [[∀pas.[σ]iN]]$ (here we assume that $[[nas]]$ and $[[pas]]$ are lists of fresh variables, 
          that is the variable capture never happens);
      \end{itemize}
    \end{multicols}
\end{algorithm}

@@

Sometimes it is convenient to construct substitution 
explicitly mapping each variable from a list (or a set)
to a type. Such substitutions are denoted as $[[iPs / pas]]$
and $[[iNs / nas]]$, where $[[iPs]]$ and $[[iNs]]$ are lists of 
the corresponding types.
\begin{definition}[Explicit Substitution]
  \hfill
  \begin{itemize}
    \item [$-$]
      Suppose that $[[nas]]$ is a list of negative type variables,
      and $[[iNs]]$ is a list of negative types of the same length.
      Then $[[iNs / nas]]$ denotes a substitution such that 
      \begin{enumerate}
        \item for $[[αi⁺ ∊ {nas}]]$, $[[ [iNs / nas] αi⁺]] = [[iNi]]$;
        \item for $[[β⁺ ∉ {nas}]]$, $[[ [iNs / nas] β⁺]] = [[β⁺]]$.
      \end{enumerate}
    \item [$+$]
      Positive explicit substitution $[[iPs / pas]]$
      is defined symmetrically.
  \end{itemize}
\end{definition}

\subsection{Declarative Subtyping}
Subtyping is one of the key mechanism of our system. 
It realizes the polymorphism: abstract $[[∀]]$ and 
$[[∃]]$ types can be used where concrete types are expected,
exactly because the subtyping relation between them.

\begin{definition} 
  \label{def:subDOne}
  \hfill

  \begin{multicols}{2}
    \ottdefnDOneNsub{}
    \columnbreak\\
    \ottdefnDOnePsup{}
  \end{multicols}

  \begin{multicols}{2}
    \ottdefnDOneNeq{}
    \columnbreak\\
    \ottdefnDOnePeq{}
  \end{multicols}
\end{definition}

The following observations about the declarative subtyping are worth noting:
\begin{itemize}
  \item \ruleref{\ottdruleDOneNVarLabel} and \ruleref{\ottdruleDOnePVarLabel}
    make the subtyping reflexive on variables (and further, on any type).
  \item \ruleref{\ottdruleDOneArrowLabel} is standard: the arrow is covariant on the
    resulting type and contravariant on the argument type.
  \item \ruleref{\ottdruleDOneShiftDLabel}  and \ruleref{\ottdruleDOneShiftULabel} are non-standard:
    the subtyping is \emph{invariant} for shifts. 
    This way, the subtyping of shifted types in one direction implies the subtyping
    in the opposite direction.
    Although this rule restricts the
    subtyping relation, it makes the system decidable.
  \item \ruleref{\ottdruleDOneForallLabel} and \ruleref{\ottdruleDOneExistsLabel} are the only
    non-algorithmic rules: the substitution for the quantified variable is
    not specified, those, these rules `drive' the subtyping relation.
\end{itemize}

In the next section, we present the sound and complete algorithm
checking whether one type is a subtype of another according to \cref{def:subDOne}.

@@

\begin{definition}[Algorithmic Types]
  \label{def:algo-types}
  \hfill\\
  \begin{multicols}{2}
    \ottgrammartabular{
      \ottuN\ottinterrule
    }

    \ottgrammartabular{
      \ottuP\ottinterrule
    }
    \columnbreak
  \end{multicols}
\end{definition}

\subsection{Fresh Variable Selection}
\label{sec:fresh-selection}
Both the subtyping and the type inference algorithm
rely on the ability to select fresh, unused variables.
For a set of variables $[[varset]]$, it is indicated as 
$[[varset are fresh]]$ in the inference rules.
We assume that the selection subroutine always succeeds and is 
deterministic. In other words, whenever it is called in 
an algorithmic inference rule, it returns the same result, 
uniquely determined by the input of this rule.

\subsection{Variable Algorithmization}
\label{sec:variable-algorithmization}

In several places of our algorithm, in particular, during
algorithmic subtyping,
we turn a declarative type into the algorithmic one
via replacing certain type variables with fresh algorithmic variables.
We call this procedure \emph{variable algorithmization}, and define it as follows.

\begin{definition}[Variable Algorithmization]
  Suppose that $[[nas]]$ is a list of negative type variables
  and $[[nuas]]$ is a list of negative algorithmic variables of the same length. 
  Then $[[ nuas/nas ]]$ is a substitution-like procedure replacing each $[[αi⁻ ∊ {nas}]]$
  in a type for $[[αî⁻ ∊ {nuas}]]$.
\end{definition}

Conversely, we have the opposite procedure turning algorithmic type variables
into declarative type variables via \emph{dealgorithmization}.

\begin{definition}[Variable Dealgorithmization]
  Suppose that $[[nuas]]$ is a list of negative algorithmic variables
  and $[[nas]]$ is a list of negative type variables of the same length. 
  Then $[[ nas/nuas ]]$ is a substitution-like procedure replacing each
  $[[αî⁻ ∊ {nuas}]]$ in a type for $[[αi⁻ ∊ {nas}]]$.
\end{definition}

@@

\begin{definition}[Algorithmic Type Context $[[Ξ]]$]
  \hfill \\
  Algorithmic type context $[[Ξ]]$ is represented by a set of 
  \emph{algorithmic} type variables ($[[α̂⁺]]$, $[[α̂⁻]]$, $[[β̂⁺]]$, \dots).
  The concatenation $[[Ξ1, Ξ2]]$ means the union of two contexts $[[Ξ1 ∪ Ξ2]]$.
\end{definition}

@@

\begin{algorithm}[Algorithmic Type Well-formedness]
  \hfill

  \begin{multicols}{2}
  \ottdefnWFATNauWF{}
  \columnbreak

  \ottdefnWFATPauWF{}
  \end{multicols}

\end{algorithm}

Algorithmic Type Context are used in the unification algorithm.
In the subtyping algorithm, 
the context needs to remember additional information.
In the subtyping context, each algorithmic variable is associated with a
context it must be instantiated in 
(i.e. the context in which the type replacing the variable must be well-formed).
This association is represented by \emph{algorithmic subtyping context} $[[Θ]]$.
\begin{definition}[Algorithmic Subtyping Context $[[Θ]]$]
  \hfill \\
  Algorithmic Subtyping Context $[[Θ]]$ is represented by a set of 
  entries of form $[[ α̂⁺[Γ] ]]$ and $[[ α̂⁻[Γ] ]]$,
  where $[[α̂⁺]]$ and $[[α̂⁻]]$ are algorithmic variables,
  and $[[Γ]]$ is a context in which they must be instantiated.
  We assume that no two entries associating the same variable
  appear in $[[Θ]]$.

  $[[dom(Θ)]]$ denotes the set of variables appearing in $[[Θ]]$:
  $[[dom(Θ)]] = \{ [[α̂±]] \mid [[α̂±[Γ] ]] \in [[Θ]] \}$.
  If $[[ α̂±[Γ] ]] \in [[Θ]]$, we denote $[[Γ]]$ as $[[Θ(α̂±)]]$.
\end{definition}

@@

Substitution that operates on algorithmic type variables is denoted as
$[[uσ]]$. It is defined as a total function from algorithmic 
type variables to \emph{declarative} types, preserving the polarity.

@@

\subsection{Equivalence and Normalization}
